---
title: "semmcci"
author: "Ivan Jacob Agaloos Pesigan"
date: "`r Sys.Date()`"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(42)
```

<!-- badges: start -->
[![CRAN Status](https://www.r-pkg.org/badges/version/semmcci)](https://cran.r-project.org/package=semmcci)
[![R-Universe Status](https://jeksterslab.r-universe.dev/badges/semmcci)](https://jeksterslab.r-universe.dev)
[![R-CMD-check](https://github.com/jeksterslab/semmcci/workflows/R-CMD-check/badge.svg)](https://github.com/jeksterslab/semmcci/actions)
[![test-coverage](https://github.com/jeksterslab/semmcci/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/jeksterslab/semmcci/actions/workflows/test-coverage.yaml)
[![codecov](https://codecov.io/gh/jeksterslab/semmcci/branch/main/graph/badge.svg?token=KVLUET3DJ6)](https://codecov.io/gh/jeksterslab/semmcci)
<!-- badges: end -->

## Installation

You can install the CRAN release of `semmcci` with:

```r
install.packages("semmcci")
```

You can install the development version of `semmcci` from
[GitHub](https://github.com/jeksterslab/semmcci) with:

```r
if (!require("remotes")) install.packages("remotes")
remotes::install_github("jeksterslab/semmcci")
```

## Documentation

See [GitHub Pages](https://jeksterslab.github.io/semmcci/index.html) for
package documentation.

## Description

In the Monte Carlo method, a sampling distribution of parameter
estimates is generated from the multivariate normal distribution using
the parameter estimates and the sampling variance-covariance matrix.
Confidence intervals for defined parameters are generated by obtaining
percentiles corresponding to 100(1 - α)% from the generated sampling
distribution, where α is the significance level.

Monte Carlo confidence intervals for free and defined parameters in
models fitted in the structural equation modeling package `lavaan` can
be generated using the `semmcci` package. The package has two main
functions, namely, `MC()` and `MCStd()`. The output of `lavaan` is
passed as the first argument to the `MC()` function to generate Monte
Carlo confidence intervals. Monte Carlo confidence intervals for the
standardized estimates can also be generated by passing the output of
the `MC()` function to the `MCStd()` function.

## Example

A common application of the Monte Carlo method
is to generate confidence intervals for the indirect effect.
In the simple mediation model,
variable `X` has an effect on variable `Y`,
through a mediating variable `M`.
This mediating or indirect effect
is a product of path coefficients from the fitted model.

```{r, message = FALSE}
library(semmcci)
library(lavaan)
```

### Data

```{r}
n <- 1000
X <- rnorm(n = n)
M <- 0.50 * X + rnorm(n = n)
Y <- 0.25 * X + 0.50 * M + rnorm(n = n)
data <- data.frame(X, M, Y)
```

### Model Specification

The indirect effect is defined by the product of the slopes
of paths `X` to `M` labeled as `a` and `M` to `Y` labeled as `b`.
In this example, we are interested in the confidence intervals of `indirect`
defined as the product of `a` and `b` using the `:=` operator
in the `lavaan` model syntax.

```{r}
model <- "
  Y ~ cp * X + b * M
  M ~ a * X
  indirect := a * b
  direct := cp
  total := cp + (a * b)
"
```

### Model Fitting

We can now fit the model using the `sem()` function from `lavaan`.

```{r}
fit <- sem(data = data, model = model)
```

### Monte Carlo Confidence Intervals

The `fit` `lavaan` object can then be passed to the `MC()` function
to generate Monte Carlo confidence intervals.

```{r}
MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
```

### Standardized Monte Carlo Confidence Intervals

Standardized Monte Carlo Confidence intervals can be generated by passing the result of the `MC()` function to `MCStd()`.

> **Note:** We recommend setting `fixed.x = FALSE` when generating standardized estimates and confidence intervals to model the variances and covariances of the predictors if they are assumed to be random.

```{r}
fit <- sem(data = data, model = model, fixed.x = FALSE)
unstd <- MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
```

```{r}
MCStd(unstd)
```
