---
title: "Example 1: The Simple Mediation Model"
author: "Ivan Jacob Agaloos Pesigan"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Example 1: The Simple Mediation Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include = FALSE
set.seed(42)
```

```{r}
#| include = FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A common application of the Monte Carlo method
is to generate confidence intervals for the indirect effect.
In the simple mediation model,
variable `X` has an effect on variable `Y`,
through a mediating variable `M`.
This mediating or indirect effect
is a product of path coefficients from the fitted model.


```{tikz}
#| echo = FALSE, 
#| fig.cap = "The Simple Mediation Model", 
#| fig.ext = 'png', 
#| cache = FALSE
\usetikzlibrary{
	er,
	arrows,
	positioning
}

\begin{tikzpicture}[
		auto,
		node distance = 20mm,
		manifest/.style = {
			rectangle,
			draw,
			thick,
			inner sep = 0pt,
			minimum width = 15mm,
			minimum height = 10mm
		},
		inv/.style = {
			rectangle,
			draw=none,
			fill=none,
			inner sep = 0pt,
			minimum width = 15mm,
			minimum height = 10mm
		},
		error/.style = {
			ellipse,
			draw,
			thick,
			inner sep = 0pt,
			minimum size = 7mm,
			align = center
		},
		mean/.style={
			regular polygon,
			regular polygon sides = 3,
			draw,
			thick,
			inner sep = 0pt,
			minimum width = 7mm,
			minimum height = 7mm
		},
		path/.style = {
			->,
			thick,
			>=stealth'
		},
		cov/.style = {
			<->,
			thick,
			>=stealth'
		},
	]
	\node[manifest] (X) {$X$};
	\node[manifest] (M) [above right = of X] {$M$};
	\node[manifest] (Y) [below right = of M] {$Y$};
	\node[error] (eM) [above = of M] {$e_{M}$};
	\node[error] (eY) [above = of Y] {$e_{Y}$};
	\draw[path] (M) to node[above] {$b$} (Y);
	\draw[path] (X) to node {$c^{\prime}$} (Y);
	\draw[path] (X) to node {$a$} (M);
	\draw[path] (eM) to node {$1$} (M);
	\draw[path] (eY) to node {$1$} (Y);
	\draw[cov] (X) to[out = 70, in = 110, looseness = 5] node[above] {$s_{X}^{2}$} (X);
	\draw[cov] (eM) to[out = -290, in = -250, looseness = 5] node[above] {$s^{2}_{e_{M}}$} (eM);
	\draw[cov] (eY) to[out = -290, in = -250, looseness = 5] node[above] {$s^{2}_{e_{Y}}$} (eY);
\end{tikzpicture}
```

```{r}
#| message = FALSE
library(semmcci)
library(lavaan)
```

## Data

```{r}
n <- 1000
X <- rnorm(n = n)
M <- 0.50 * X + rnorm(n = n)
Y <- 0.25 * X + 0.50 * M + rnorm(n = n)
data <- data.frame(X, M, Y)
```

## Model Specification

The indirect effect is defined by the product of the slopes
of paths `X` to `M` labeled as `a` and `M` to `Y` labeled as `b`.
In this example, we are interested in the confidence intervals of `indirect`
defined as the product of `a` and `b` using the `:=` operator
in the `lavaan` model syntax.

```{r}
model <- "
  Y ~ cp * X + b * M
  M ~ a * X
  indirect := a * b
  direct := cp
  total := cp + (a * b)
"
```

## Model Fitting

We can now fit the model using the `sem()` function from `lavaan`.

```{r}
fit <- sem(data = data, model = model)
```

## Monte Carlo Confidence Intervals

The `fit` `lavaan` object can then be passed to the `MC()` function
to generate Monte Carlo confidence intervals.

```{r}
MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
```

## Standardized Monte Carlo Confidence Intervals

Standardized Monte Carlo Confidence intervals can be generated by passing the result of the `MC()` function to `MCStd()`.

> **Note:** We recommend setting `fixed.x = FALSE` when generating standardized estimates and confidence intervals to model the variances and covariances of the predictors if they are assumed to be random.

> **Note:** We recommend setting `par = TRUE` and `ncores` to the number of cores to use when generating standardized estimates and confidence intervals in order to speed up the process.

```{r}
fit <- sem(data = data, model = model, fixed.x = FALSE)
unstd <- MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
```

```{r}
#| eval = FALSE
MCStd(unstd, par = TRUE, ncores = parallel::detectCores())
```

```{r}
#| echo = FALSE
MCStd(unstd)
```

## Standardized Monte Carlo Confidence Intervals - An Alternative Approach

In this example, confidence intervals for the standardized indirect effect
are generated by specifying the standardized indirect effect
as a derived parameter using the `:=` operator.
The standardized indirect effect in a simple mediation model
involves paths $a$ and $b$,
and the standard deviations of $X$ and $Y$.
It is given by

\begin{equation}
  a b
  \frac{s_X}{s_Y}
\end{equation}

\noindent where

\begin{equation}
  s_X
  =
  \sqrt{
    s_{X}^{2}
  }
\end{equation}

\noindent and

\begin{equation}
  s_Y
  =
  \sqrt{
    c^{\prime 2} s_{X}^{2}
    + a^2 b^2 s_{X}^{2}
    + b^2 s_{e_{M}}^{2}
    + 2 c^{\prime} b a s_{X}^{2}
    + s_{e_{Y}}^{2}
  }
\end{equation}

\noindent where

\noindent $s_{e_{Y}}^{2}$ and $s_{e_{M}}^{2}$ are the residual variances
in the regression equations.

The standardized indirect effect can be defined using the `:=` operator
and the named parameters in the model.

```{r}
model <- "
  Y ~ cp * X + b * M
  M ~ a * X
  X ~~ s2_X * X
  M ~~ s2_eM * M
  Y ~~ s2_eY * Y
  indirect_std := a * b * (sqrt(s2_X) / sqrt(cp^2 * s2_X + a^2 * b^2 * s2_X + b^2 * s2_eM + 2 * cp * b * a * s2_X + s2_eY))
"
```

```{r}
fit <- sem(data = data, model = model, fixed.x = FALSE)
```

The row `indirect_std` corresponds to the confidence intervals for the standardized indirect effect.

```{r}
MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
```
