---
title: "Example 4.2: The Simple Mediation Model with Missing Data (MI)"
author: "Ivan Jacob Agaloos Pesigan"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Example 4.2: The Simple Mediation Model with Missing Data (MI)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include = FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To cite `semmcci` in publications, please use: 

Pesigan, I. J. A., & Cheung, S. F. (2023).
Monte Carlo confidence intervals for the indirect effect with missing data.
*Behavior Research Methods*.
https://doi.org/10.3758/s13428-023-02114-4

```{r}
#| include = FALSE
set.seed(42)
```

In this example,
the Monte Carlo method is used to generate confidence intervals
for the indirect effect in a simple mediation model with missing data
where
variable `X` has an effect on variable `Y`,
through a mediating variable `M`
with multiply imputed data sets.

```{r}
#| message = FALSE
library(semmcci)
library(lavaan)
```

## Data

```{r}
#| echo = FALSE
n <- 1000
a <- 0.50
b <- 0.50
cp <- 0.25
s2_em <- 1 - a^2
s2_ey <- 1 - cp^2 - a^2 * b^2 - b^2 * s2_em - 2 * cp * a * b
em <- rnorm(n = n, mean = 0, sd = sqrt(s2_em))
ey <- rnorm(n = n, mean = 0, sd = sqrt(s2_ey))
X <- rnorm(n = n)
M <- a * X + em
Y <- cp * X + b * M + ey
df <- data.frame(X, M, Y)

# Create data set with missing values.

miss <- sample(seq_len(dim(df)[1]), 300)
df[miss[1:100], "X"] <- NA
df[miss[101:200], "M"] <- NA
df[miss[201:300], "Y"] <- NA
```

```{r}
summary(df)
```

## Model Specification

The indirect effect is defined by the product of the slopes
of paths `X` to `M` labeled as `a` and `M` to `Y` labeled as `b`.
In this example, we are interested in the confidence intervals of `indirect`
defined as the product of `a` and `b` using the `:=` operator
in the `lavaan` model syntax.

```{r}
model <- "
  Y ~ cp * X + b * M
  M ~ a * X
  indirect := a * b
  direct := cp
  total := cp + (a * b)
"
```

## Model Fitting

We can now fit the model using the `sem()` function from `lavaan`.
We do not need to deal with missing values in this stage.
We also set `fixed.x = FALSE` to estimate variances and covariances
of exogenous variables.

```{r}
fit <- sem(data = df, model = model, fixed.x = FALSE)
```

## Monte Carlo Confidence Intervals

The `fit` `lavaan` object can then be passed to the `MCMI()` function from `semmcci`
to generate Monte Carlo confidence intervals
using multiple imputation.
Additional arguments to the function used for multiple
imputation can be supplied
to override the default options.
For example,
the argument `m` can be used to increase the number of imputations
from the default 5.
See the `...` argument in the `MCMI()` function
and the arguments in `mice::mice()`.

```{r}
#| include = FALSE
set.seed(42)
```

```{r}
MCMI(fit, R = 20000L, alpha = 0.05, m = 100)
```

`MCMI()` automatically generates multiply imputed data sets
using `mice::mice()`.
Alternatively, a list of multiply imputed data sets can be supplied
using the `imp` argument.

```{r}
#| include = FALSE
set.seed(42)
```

```{r}
#| echo = FALSE
imp <- mice::complete(
  mice::mice(
    data = df,
    print = FALSE,
    m = 100
  ),
  action = "all"
)
```

```{r}
length(imp)
head(imp[[1]])
```

```{r}
MCMI(fit, R = 20000L, alpha = 0.05, imp = imp)
```

## Standardized Monte Carlo Confidence Intervals

Standardized Monte Carlo Confidence intervals can be generated by passing the result of the `MCMI()` function to the `MCStd()` function.

```{r}
fit <- sem(data = df, model = model, fixed.x = FALSE)
unstd <- MCMI(fit, R = 20000L, alpha = 0.05, m = 100)
```

```{r}
MCStd(unstd, alpha = 0.05)
```
