---
title: "Example 1: The Simple Mediation Model"
author: "Ivan Jacob Agaloos Pesigan"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Example 1: The Simple Mediation Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is an example of a simple mediation model
where `X` is the predictor, `M` is the mediator,
and `Y` is the dependent variable.

```{r, message = FALSE}
library(semmcci)
library(lavaan)
```

## Data

```{r}
n <- 1000
X <- rnorm(n = n)
M <- 0.50 * X + rnorm(n = n)
Y <- 0.25 * X + 0.50 * M + rnorm(n = n)
data <- data.frame(X, M, Y)
```

## Model Specification

The indirect effect is defined by the product of the slopes
of paths `X` to `M` labeled as `a` and `M` to `Y` labeled as `b`.
In this example, we are interested in the confidence intervals of `ab`
defined as the product of `a` and `b` using the `:=` operator
in the `lavaan` model syntax.

```{r}
model <- "
  Y ~ cp * X + b * M
  M ~ a * X
  indirect := a * b
  direct := cp
  total := cp + (a * b)
"
```

## Model Fitting

We can now fit the model using the `sem()` function from `lavaan`.

```{r}
fit <- sem(data = data, model = model)
```

## Monte Carlo Confidence Intervals

The `fit` `lavaan` object can then be passed to the `MC()` function from `semmcci`
to generate Monte Carlo confidence intervals.

```{r}
MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
```

## Standardized Monte Carlo Confidence Intervals

Standardized Monte Carlo Confidence intervals can be generated by passing the result of the `MC()` function to `MCStd()`.

> **Note:** We recommend setting `fixed.x = FALSE` when generating standardized estimates and confidence intervals to model the variance/s of the predictor/s.

```{r}
fit <- sem(data = data, model = model, fixed.x = FALSE)
unstd <- MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
MCStd(unstd)
```

## Standardized Monte Carlo Confidence Intervals - An Alternative Approach

In this example, confidence intervals for the standardized indirect effect
are generated by specifying the standardized indirect effect
as a derived parameter using the `:=` operator.
The standardized indirect effect in a simple mediation model
involves paths $a$, and $b$,
and the standard deviations of $X$ and $Y$.
It is given by

\begin{equation}
  a b
  \frac{s_X}{s_Y}
\end{equation}

\noindent where

\begin{equation}
  s_X
  =
  \sqrt{
    s_{X}^{2}
  }
\end{equation}

\noindent and

\begin{equation}
  s_Y
  =
  \sqrt{
    c^{\prime 2} s_{X}^{2}
    + a^2 b^2 s_{X}^{2}
    + b^2 s_{e_{M}}^{2}
    + 2 c^{\prime} b a s_{X}^{2}
    + s_{e_{Y}}^{2}
  }
\end{equation}

\noindent where

\noindent $s_{e_{Y}}^{2}$ and $s_{e_{M}}^{2}$ are the residual variances
in the regression equations.

The standardized indirect effect can be defined using the `:=` operator
and the named parameters in the model.

```{r}
model <- "
  Y ~ cp * X + b * M
  M ~ a * X
  X ~~ var_X * X
  M ~~ var_eM * M
  Y ~~ var_eY * Y
  indirect_std := a * b * (sqrt(var_X) / sqrt(cp^2 * var_X + a^2 * b^2 * var_X + b^2 * var_eM + 2 * cp * b * a * var_X + var_eY))
"
```

```{r}
fit <- sem(data = data, model = model, fixed.x = FALSE)
```

The row `indirect_std` corresponds to the confidence intervals for the defined parameter.

```{r}
MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
```
