---
title: "Example 4: The Simple Mediation Model with Missing Data"
author: "Ivan Jacob Agaloos Pesigan"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Example 4: The Simple Mediation Model with Missing Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include = FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| include = FALSE
set.seed(42)
```

In this example,
the Monte Carlo method is used to generate confidence intervals
for the indirect effect in a simple mediation model with missing data
where
variable `X` has an effect on variable `Y`,
through a mediating variable `M`.

```{r}
#| message = FALSE
library(semmcci)
library(lavaan)
```

## Data

```{r}
n <- 1000
a <- 0.50
b <- 0.50
cp <- 0.25
s2_em <- 1 - a^2
s2_ey <- 1 - cp^2 - a^2 * b^2 - b^2 * s2_em - 2 * cp * a * b
em <- rnorm(n = n, mean = 0, sd = sqrt(s2_em))
ey <- rnorm(n = n, mean = 0, sd = sqrt(s2_ey))
X <- rnorm(n = n)
M <- a * X + em
Y <- cp * X + b * M + ey
df <- data.frame(X, M, Y)

# Create data set with missing values.

miss <- sample(seq_len(dim(df)[1]), 300)
df[miss[1:100], "X"] <- NA
df[miss[101:200], "M"] <- NA
df[miss[201:300], "Y"] <- NA
```

## Model Specification

The indirect effect is defined by the product of the slopes
of paths `X` to `M` labeled as `a` and `M` to `Y` labeled as `b`.
In this example, we are interested in the confidence intervals of `indirect`
defined as the product of `a` and `b` using the `:=` operator
in the `lavaan` model syntax.

```{r}
model <- "
  Y ~ cp * X + b * M
  M ~ a * X
  indirect := a * b
  direct := cp
  total := cp + (a * b)
"
```

## Model Fitting

We can now fit the model using the `sem()` function from `lavaan`.
We are using `missing = "fiml"` to handle missing data in `lavaan`.
Since there are missing values in `x`, we also set `fixed.x = FALSE`.

```{r}
fit <- sem(data = df, model = model, missing = "fiml", fixed.x = FALSE)
```

## Monte Carlo Confidence Intervals

The `fit` `lavaan` object can then be passed to the `MC()` function from `semmcci`
to generate Monte Carlo confidence intervals.

```{r}
MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
```

## Standardized Monte Carlo Confidence Intervals

Standardized Monte Carlo Confidence intervals can be generated by passing the result of the `MC()` function to the `MCStd()` function.

```{r}
fit <- sem(data = df, model = model, missing = "fiml", fixed.x = FALSE)
unstd <- MC(fit, R = 20000L, alpha = c(0.001, 0.01, 0.05))
```

```{r}
MCStd(unstd)
```
