---
title: "semmcci"
author: "Ivan Jacob Agaloos Pesigan"
date: "`r Sys.Date()`"
output: github_document
bibliography: "vignettes.bib"
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
nocite: |
  @*
---

<!-- README.md is generated from .setup/readme/README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Make Project](https://github.com/jeksterslab/semmcci/actions/workflows/make.yml/badge.svg)](https://github.com/jeksterslab/semmcci/actions/workflows/make.yml)
[![R-CMD-check](https://github.com/jeksterslab/semmcci/actions/workflows/check-full.yml/badge.svg)](https://github.com/jeksterslab/semmcci/actions/workflows/check-full.yml)
[![R Package Test Coverage](https://github.com/jeksterslab/semmcci/actions/workflows/test-coverage.yml/badge.svg)](https://github.com/jeksterslab/semmcci/actions/workflows/test-coverage.yml)
[![Lint R Package](https://github.com/jeksterslab/semmcci/actions/workflows/lint.yml/badge.svg)](https://github.com/jeksterslab/semmcci/actions/workflows/lint.yml)
[![Package Website (GitHub Pages)](https://github.com/jeksterslab/semmcci/actions/workflows/pkgdown-gh-pages.yml/badge.svg)](https://github.com/jeksterslab/semmcci/actions/workflows/pkgdown-gh-pages.yml)
[![Compile LaTeX](https://github.com/jeksterslab/semmcci/actions/workflows/latex.yml/badge.svg)](https://github.com/jeksterslab/semmcci/actions/workflows/latex.yml)
[![Shell Check](https://github.com/jeksterslab/semmcci/actions/workflows/shellcheck.yml/badge.svg)](https://github.com/jeksterslab/semmcci/actions/workflows/shellcheck.yml)
[![pages-build-deployment](https://github.com/jeksterslab/semmcci/actions/workflows/pages/pages-build-deployment/badge.svg)](https://github.com/jeksterslab/semmcci/actions/workflows/pages/pages-build-deployment)
[![codecov](https://codecov.io/gh/jeksterslab/semmcci/branch/main/graph/badge.svg?token=KVLUET3DJ6)](https://codecov.io/gh/jeksterslab/semmcci)
<!-- badges: end -->

## Installation

You can install the CRAN release of `semmcci` with:

```r
install.packages("semmcci")
```

You can install the development version of `semmcci` from
[GitHub](https://github.com/jeksterslab/semmcci) with:

```r
if (!require("remotes")) install.packages("remotes")
remotes::install_github("jeksterslab/semmcci")
```

## Description

In the Monte Carlo method, a sampling distribution of parameter
estimates is generated from the multivariate normal distribution using
the parameter estimates and the sampling variance-covariance matrix.
Confidence intervals for defined parameters are generated by obtaining
percentiles corresponding to 100(1 - α)% from the generated sampling
distribution, where α is the significance level.

Monte Carlo confidence intervals for free and defined parameters in
models fitted in the structural equation modeling package `lavaan` can
be generated using the `semmcci` package. The package has three main
functions, namely, `MC()`, `MCMI()`, and `MCStd()`. The output of `lavaan` is
passed as the first argument to the `MC()` function or the `MCMI()` function
to generate Monte Carlo confidence intervals. Monte Carlo confidence intervals for the
standardized estimates can also be generated by passing the output of
the `MC()` function or the `MCMI()` function to the `MCStd()` function.
A description of the package and code examples
are presented in Pesigan and Cheung (2023: https://doi.org/10.3758/s13428-023-02114-4).

## Example

A common application of the Monte Carlo method
is to generate confidence intervals for the indirect effect.
In the simple mediation model,
variable `X` has an effect on variable `Y`,
through a mediating variable `M`.
This mediating or indirect effect
is a product of path coefficients from the fitted model.

```{r, message = FALSE}
library(semmcci)
library(lavaan)
```

### Data

```{r}
#| echo = FALSE
n <- 1000
a <- 0.50
b <- 0.50
cp <- 0.25
sigmasqem <- 1 - a^2
sigmasqey <- 1 - cp^2 - a^2 * b^2 - b^2 * (1 - a^2) - 2 * cp * a * b
em <- rnorm(n = n, mean = 0, sd = sqrt(sigmasqem))
ey <- rnorm(n = n, mean = 0, sd = sqrt(sigmasqey))
X <- rnorm(n = n)
M <- a * X + em
Y <- cp * X + b * M + ey
df <- data.frame(X, M, Y)

# Create data set with missing values.

miss <- sample(seq_len(dim(df)[1]), 300)
df[miss[1:100], "X"] <- NA
df[miss[101:200], "M"] <- NA
df[miss[201:300], "Y"] <- NA
```

```{r}
summary(df)
```

### Model Specification

The indirect effect is defined by the product of the slopes
of paths `X` to `M` labeled as `a` and `M` to `Y` labeled as `b`.
In this example, we are interested in the confidence intervals of `indirect`
defined as the product of `a` and `b` using the `:=` operator
in the `lavaan` model syntax.

```{r}
model <- "
  Y ~ cp * X + b * M
  M ~ a * X
  X ~~ X
  indirect := a * b
  direct := cp
  total := cp + (a * b)
"
```

### Monte Carlo Confidence Intervals

We can now fit the model using the `sem()` function from `lavaan`.
We use full-information maximum likelihood to deal with missing values.

```{r}
fit <- sem(data = df, model = model, missing = "fiml")
```

The `fit` `lavaan` object can then be passed to the `MC()` function
to generate Monte Carlo confidence intervals.

```{r}
mc <- MC(fit, R = 20000L, alpha = 0.05)
mc
```

### Monte Carlo Confidence Intervals - Multiple Imputation

The `MCMI()` function can be used to handle missing values using multiple imputation.
The `MCMI()` accepts the output of
`mice::mice()`,
`Amelia::amelia()`,
or a list of multiply imputed data sets.
In this example, we impute multivariate missing data under the normal model.

```{r}
mi <- mice::mice(
  df,
  method = "norm",
  m = 100,
  print = FALSE,
  seed = 42
)
```

We fit the model using lavaan using the default listwise deletion.

```{r}
fit <- sem(data = df, model = model)
```

The `fit` `lavaan` object and `mi` object can then be passed to the `MCMI()` function
to generate Monte Carlo confidence intervals.

```{r}
mcmi <- MCMI(fit, mi = mi, R = 20000L, alpha = 0.05, seed = 42)
mcmi
```

### Standardized Monte Carlo Confidence Intervals

Standardized Monte Carlo Confidence intervals can be generated
by passing the result
of the `MC()` function or the `MCMI()` function to `MCStd()`.

```{r}
MCStd(mc, alpha = 0.05)
```

```{r}
MCStd(mcmi, alpha = 0.05)
```

## Citation

To cite `semmcci` in publications, please use: 

Pesigan, I. J. A., & Cheung, S. F. (2023).
Monte Carlo confidence intervals for the indirect effect with missing data.
*Behavior Research Methods*.
https://doi.org/10.3758/s13428-023-02114-4

## Documentation

See [GitHub Pages](https://jeksterslab.github.io/semmcci/index.html) for
package documentation.
